---
title: "**Assignment 3 Naive Bayes**"
author: "wliu16"
date: "2022-10-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```

```{r, include= FALSE}
#load the data and observe the data
rm(list=ls())
setwd("/Users/tammyliu/Desktop/R/64060-003/Assignment_3")
bank.df <- read.csv("UniversalBank.csv")
```

```{r, include=FALSE}
library(caret)
library(dplyr)
library(reshape2) #melting and casting
library(reshape)
library(gmodels)
library(pROC)
library(e1071)
```


```{r}
#select, rename and clean the data
bank_data <- bank.df[, c(10, 13, 14)]
bank_data$Online <- as.factor(bank_data$Online)
bank_data$CC <- as.factor(bank_data$CreditCard)
bank_data$Loan <- as.factor(bank_data$Personal.Loan)
bank_data <- bank_data[, c(-1, -3)]

#Data partition with training: validation = 6:4
set.seed(123)
index_train <- createDataPartition(bank_data$Loan, 
                                   p = 0.6,
                                   list = FALSE)
train_data <- bank_data[index_train, ]
validate_data <- bank_data[-index_train, ]

```

A. Pivot table of Online, CC and Loan for training data
```{r}
#molted <- melt(train_data, id = c("Online"))
#tibble(molted)
#x.cast <- dcast(molted, Online~variable, fun.aggregate = length)
#x.cast
```
```{r}
xtab <- ftable(xtabs(~CC+Loan+Online, data = train_data))
xtab
```

B. P(Loan=1 | CC=1, Online=1) = 51/(467+51) = 9.85%

C. Create two separate pivot tables
```{r}
melt.x1 <- melt(train_data, id = "Loan", variable = "Online")
cast.x1 <- dcast(melt.x1, Loan~Online, fun.aggregate = length)
cast.x1

ftable(train_data[, c(3,1)]) #Loan as row vs Online as column
ftable(train_data[, c(3,2)]) #Loan as row vs CC as column
```

D. Compute P(A | B)
```{r}
ftable(train_data[, c(2,3)]) #CC | Loan
ftable(train_data[, c(1,3)]) #Online | Loan
ftable(train_data[, 3]) #Loan distribution 
```

i. P(CC = 1 | Loan =1) = 84/(204+84) = 0.2917
ii. P(Online =1 | Loan =1) = 176/(176+112) = 0.6111
iii. P(Loan =1) = 288/(288+2712) = 0.096
iv. P(CC=1| Loan =0) = 777/(1935+777) = 0.2865
v. P(Online =1 | Loan =0) = 1611/(1611+1101) = 0.5940
vi. P(Loan = 0) = 2712/(2712+288) = 0.904

E. Naive Bayes probability
P(Loan =1 | CC=1, Online =1) = 
P(CC = 1 | Loan =1)*P(Online =1|Loan =1)*P(Loan =1)/
(P(CC = 1 | Loan =1)*P(Online =1|Loan =1)*P(Loan =1)+
P(CC=1| Loan =0)*P(Online =1 | Loan =0)*P(Loan = 0))=
0.2971*0.6111*0.096/(0.2971*0.6111*0.096+0.2865*0.5940*0.904)= 0.1018 =10.18%

F. Naive Bayes result of 10.18% is higher than pivot table calculation of 9.84%

G. Run NB
```{r}
result <- naiveBayes(Loan~Online+CC, data = train_data)
result

```
The entries needed for NB calculation: 
P(loan) 0.904 and 0.096
P(Loan =0 | Online =1) = 0.5940
P(Loan =1 | Online =1) = 0.6111
P(Loan =0 | CC =1) = 0.2865
P(Loan =1 | CC =1) = 0.2917
