---
title: "Assignment_5"
author: "wliu16"
date: "2022-12-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(cluster)
library(factoextra)
library(stats)
library(corrplot)
library(fpc)
```

**Data preprocess**
```{r}
cereal <- read.csv("Cereals.csv")
cereal <- na.omit(cereal) # 3 NA records removed, 74 records in total
summary(cereal) #original data summary
head(cereal)  #original data snapshot

corrmatrix <- cor(cereal[, 4:16])
corrplot(corrmatrix, method = 'number')

#data scaling 
df <- cereal[, 4:16]
df_scaled<- scale(df)
rownames(df_scaled) <- cereal[, 1] #create new dataframe with only numerical data
```
<br>  
Early observations on the univariate data: <br>  there might be outliers on high
or low ends <br>  
protein: outliers on max<br>  
fat: outliers on max<br>  
sodium: outliers on min<br>  
fiber: outliers on max<br>  
potass: outliers on max<br>  
vitamins: outliers on max<br>  
rating: vitamins: outliers on max<br>  
<br>  
**Hierarchical clustering**
```{r}
d <- dist(df_scaled, method = "euclidean")
hc1 <- agnes(d, method = "complete")
hc2 <- agnes(d, method = "single")
hc3 <- agnes(d, method = "average")
hc4 <- agnes(d, method = "ward")

print(hc1$ac)
print(hc2$ac)
print(hc3$ac)
print(hc4$ac)
```
```{r}
pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram of agnes ward method")
rect.hclust(hc4, k = 5, border = 1:5)
```
```{r}
#comparison between complete and ward using hclust 
hc_comp <- hclust(d, method = "complete")
hc_w <- hclust(d, method = "ward.D")
plot(hc_comp, cex = 0.6, hang = -1) #complete 
plot(hc_w, cex = 0.6, hang = -1) #ward
```

<br>  
**Conclusion**<br>  
(1) The best method is ward method as the agglomerative coefficient is max 
among the four methods. Complete method is also a good alternative. 

If we use hclust to plot, complete and ward both show clean plots. Complete shows
5 clusters and ward shows 4. For this practice, we will go with ward method as 
the best method. <br>  

(2) I would choose 5 as the **number of clusters**. This is based on visual 
examination to the ward method and complete method. <br>  

```{r}
#test plot of Euclidean distance vs no of clusters optimal no of clusters

```


**Check on stability of clusters**
```{r}
set.seed(111)

train_index_c <- createDataPartition(df$rating, p= 0.6, list = FALSE)
validate_c <- df[- train_index_c, ]  # 40% as validation
train_c <- df[train_index_c, ] # 60% as training and testing

validate_c <- scale(validate_c)
train_c <- scale(train_c)

```


```{r}
d2 <- dist(train_c, method = "euclidean")
d3 <- dist(validate_c, method = "euclidean")

hc_ward_train <- agnes(d2, method = "ward")
hc_ward_validate <- agnes(d3, method = "ward")

```

```{r}
k_t <- kmeans (d2, centers = 5, nstart = 25)
k_t$centers
fviz_cluster(k_t, data = d2)
```
```{r}
k_v <- kmeans (d3, centers = 5, nstart = 25)
k_v$centers
fviz_cluster(k_v, data = d3)
```

```{r}
fviz_nbclust(df_scaled, kmeans, method='silhouette')
fviz_nbclust(df, kmeans, method = "wss")
# k = 3 and k = 8 are suggested
```



**Recommendation to Elementary school**
Should the data be normalized? <br>  
The data should not be normalized as we need to use units to filter healthy cereal. 

The standard for healthy cereal should be full of nutrients (fiber) compared
to other cereals which are simply tasty. Vitamins and minerals are
also something nice to have. According to the official guideline from FDA, 
cereals have to contain three-fourth ounces of whole grains and no more than 
1 gram of saturated fat, 230 milligrams of sodium and 2.5 grams of added sugars 
in order to be considered as healthy. <br>  
Reference: https://www.cnbc.com/2022/10/11/fda-redefined-healthy-these-7-cereals-do-not-qualify.html


