---
title: "Assignment_1"
author: "wliu16"
date: "2022-09-15"
output:
  pdf_document: default
  html_document: default
---

```{r} 
setwd("~/Desktop/R/64060/Assignment_1")

health_data <- read.table("500_Person_Gender_Height_Weight_Index.csv", TRUE, ",")
str(health_data)
summary(health_data)
head(health_data)
hist(health_data$Height, col = 'red')
```
Comments: There are over 50 individuals between 180-190cm in height which is the most popular height group in this dataset

```{r}
library(tidyverse)

table(health_data$Index)

health_data_arrange1 <- health_data %>% arrange(Height)
View(health_data_arrange1)

health_data_filter1<- health_data %>% filter(Index == 3)
View(health_data_filter1)
```
Comemnts: 
count the number of individuals by fitness levels 1-5 while 1 as extra weak, 3 as normal and 5 as extra obese
arrange a new dataframe by Height column on acsending order
filter a new dataframe by fitness level is normal (Index = 3) 

```{r}
Height_M <- health_data$Height/100
BMI <- health_data$Weight/(Height_M^2)
head(BMI)
```
Comments: convert the height to meters, calculate and print first 6 rows of BMI

```{r}
par(mfrow = c(1,2))
counts_h <- table(health_data$Height)
counts_w <- table(health_data$Weight)
barplot(counts_h, main = "Height Distribution", xlab= "Height(cm)")
barplot(counts_w, main = "Weight Distribution", xlab= "Weight(kg)")
```

Comments: plots two bargraphs on the distribution of height and weight. We can see most height is more concentrated towards 180-190cm while weight is more spread out evenly between 50-150 kg.

```{r}
par(mfrow = c(1,2))
plot(health_data$Weight, health_data$Height, main = "Height vs Weight",
     xlab = "Weight", ylab= "Height", pch=19)
plot(health_data$Weight, BMI, main = "BMI vs Weight",
     xlab = "Weight", ylab= "BMI",pch =18)
```

Comments: There is no obvious correlation between height and weight but BMI is positively correlated with weight. The greater the weight is, the higher BMI is. 